### 2022-01-03 input/proposal.tex
address comments up to IIID

\begin{proof}
  %% knowing that b_p delta is sufficient to say that processes will
  %% receive delta or detect blocking
  Assuming a chain of delivery $\pi_{xz} = [x, \ldots, z]$ of
  $\alpha_x$, we must prove that whatever the possible states of
  \nodes that belong to this chain, \TODO{WIP}
  \begin{asparadesc}
  \item [In the same partition:] $\forall y \in \pi_{xz}: \alpha_x =
    \last D_y$, and assuming $d_x(\delta_x)$, then with forwarding of
    staleness (Definition~\ref{def:fs}), $f_{\pi_{xz}[k]}(\delta_x)
    \implies \eventually d_{\pi_{xz}[k+1]}(\delta_x)$ except if
    $\delta_x \in D_{\pi_{xz}[k+1]}$ which means that $\pi_{xz}[k+1]$
    already delivered and forwarded $\delta_x$ following another such
    a chain of delivery. Therefore, $z$ eventually receives, hence
    delivers $\delta_x$.

  \item[ (\ref{lem:detectorB})] Assuming a chain of delivery
  $\pi_{xz} = [x, \ldots, z]$ of $\alpha_x$ with at least
  $\alpha_x = \last D_z$, and assuming $d_x(\delta_x)$, then by (A),
  the staleness notification reaches a \process $ y = \pi_{xz}[k]$. By
  Definition~\ref{def:fs}, the forwarding stops when $y$ already
  delivered $\delta_x$ ($\delta_x \in D_y$ covered by (A)) or
  delivered a better message $\alpha_s = \last D_y$ with
  $\alpha_s \neq \alpha_x$. With $y'=\pi_{xz}[k+1]$, forwarding of
  best (Definition~\ref{def:fbplus}) implies three possible outcomes:
  \begin{enumerate}[(i)]
  \item $\alpha_{s'} = \last D_{y'}$ with $\alpha_{s'} \neq \alpha_x$
    hence $y' \iff y$ which leaves two possible outcomes:
  \item
    $r_{y'}(\alpha_s) \wedge \alpha_x = \last D_{y'} \wedge \alpha_x <
    \alpha_s$ which means that a shorter path of delivery $\pi_{xz}'$
    exists that either forwards appropriate staleness notifications
    (A) or detects its possible blocking:
  \item
    $r_{y'}(\alpha_s) \wedge \delta_s \in D_{y'} \wedge \alpha_x =
    \last D_{y'}$ which detects a possible blocking of
    $\delta_x$. Without global knowledge, $y'$ assumes it belongs to
    the shortest and only path of delivery of $\alpha_x$
    ($\pi_{xy'} \subseteq \Pi_{xz}$) thus it does not further delegate
    the detection to another \process.
  \end{enumerate}
  \end{asparadesc}
  \vspace{-1.5em} %% ugly vspace but necessary for enumerate+qed
  %
  %% (i) If every \process (such as $a$) starting from the source
  %% delivers and forwards a removal notification $\delta_X$ when their
  %% last delivery is $\alpha_X$, then every such \process eventually
  %% delivers the removal notification $\delta_X$ except \processes (such
  %% as $b$) that delivered $\alpha_X$ from a \process that delivered a
  %% message from another partition $\alpha_Y$ since then.
  %
  %% These exceptions eventually receive and deliver $\alpha_Y$ since
  %% $\alpha_Y$ is better than $\alpha_X$ through this path, except if
  %% they already received and delivered the removal notification of
  %% $P_Y$ (such as $c$). These \processes may never receive hence
  %% deliver $\delta_X$, and may never receive hence deliver a better
  %% message than $\alpha_X$. They need an additional mechanism to
  %% eventually purge $\alpha_X$ that cannot rely on the eventual purging
  %% of $P_Y$ at \processes like $b$, to avoid deadlocks.
  %
  %% (ii) Assuming that every \process keeps an history of its past
  %% deliveries, \processes (such as $c$) can detect the inconsistency,
  %% since they receive from their parent an already deleted partition
  %% $P_Y$. This notification means that the parent discards any
  %% $\delta_Z$ with $\alpha_Z^z < \alpha_Y^y$, and most importantly, if
  %% $\delta_X$ exists, it discards it. To ensure the purge of stale
  %% messages, a detecting \process must assume the worst case that such
  %% $\delta_X$ exists, and send another kind of message, noted $\Delta$,
  %% that notify the possible removal of $P_X$.
  %
  %% (iii) A \process (such as $d$) whose last delivery is $\alpha_X$,
  %% but whose parent is neither inconsistent (like $b$) nor receiving
  %% $\delta_X$ (like $a$), eventually receives $\Delta$ from a detecting
  %% \process, either directly or transitively, for such a child \process
  %% ($d$) trusts the possible removal of $P_X$ by delivering and
  %% forwarding such $\Delta$. $\Delta$ suffers from identical blocking
  %% conditions (between $b$ and $c$) than $\delta$, leading to the same
  %% solutions of detection and propagation of $\Delta$. Eventually,
  %% every \process whose last delivery is $\alpha_X$ receives and
  %% delivers either a better $\alpha_Z$, or $\delta_X$, or $\Delta_X$.
\end{proof}

As a second step, detecting \processes must proactively purge the
system from their forwarded notifications. For instance in
Figure~\ref{fig:problemD}, $c$ could broadcast $\delta_a$ in order to
acknowledge the staleness of its own $\alpha_a^3$ and inform
neighboring \processes that delivered it as well. Unfortunately, proof
of Lemma~\ref{lem:detector} shows that there exists false positive in
the blocking detection: $\delta_x$ may not exist but $y'$ still
receives $\alpha_s$ after the delivery of $\delta_s$ from the path it
delivered $\alpha_x$. Figure~\ref{fig:undoproblem} highlights this
behavior. In Figure~\ref{fig:undoproblemB}, $a$ does not delete its
partition. In Figure~\ref{fig:undoproblemD} and because of scoped
broadcast, $c$ received the same series of messages as in
Figure~\ref{fig:problemD}. Yet, $c$ must assume the existence of
$\delta_a$ and act accordingly by forcing the staleness of its best
delivered message and disseminate this information to its neighbors.
To avoid flooding the system with false positive staleness
notifications, we propose to reduce the scope of staleness
notifications by including \TODO{only downstream \processes}. In
Figure~\ref{fig:undoproblemD}, false positives would not affect $a$
and $b$.

@@

Algorithm~\ref{algo:ascast} shows the instructions of our
implementation called \NAME that enables dynamic consistent
partitioning. It implements forwarding of best (see
Line~\ref{line:fb}) and echos (see Line~\ref{line:echo}). To implement
forwarding of staleness downstream as stated in
Definition~\ref{def:fsplus},
\begin{inparaenum}[(A)]
\item each \process maintains a vector of versions that associates the
  respective known local counter of each known source, or has-been
  source. It constitutes a summary of known progress of other
  \processes; 
\item each notification message $\alpha$ carries the list of
  identifier and counter of each \node that forwarded it. In the worst
  case, both these structures grow linearly with the number of
  \processes in the system $\mathcal{O}(V)$. Nevertheless, following
  the principles of scoped broadcast, we expect that \processes only
  acknowledge a small subset of sources and messages;
\item each staleness notifications $\delta$ only carry the identifier
  and counter of the \process~--~source or detector~--~that generated
  it. Only downstream \processes may deliver such message, since they
  carry the identifier of the generator. 
\end{inparaenum}
To implement the detection as stated in Lemma~\ref{lem:detector}, each
\process only requires to know the direct parent of its best delivered
message which is already included in the piggybacked path of this
message. Receiving a message known to be stale from this parent
triggers the generation of staleness notifications that force
downstream \processes to reevaluate the partition they are in (see
Line~\ref{line:detect}).

@@

By reusing \NAME's default behavior, Algorithm~\ref{algo:edges}
enables dynamic consistent partitioning in dynamic networks subject to
physical partitioning where \processes can join, leave, or crash at
any time. Adding a \process is equivalent to add as many edges as
necessary. Since adding an edge may improve shortest paths, the
protocol triggers a competition using echos of
Definition~\ref{def:fbplus}.  Removing a \process is equivalent to
removing all its edges. Removing an edge between two \processes may
invalidate the shortest path of one of involved \processes plus
downstream \processes, or impair the propagation of staleness
notifications. Therefore, the protocol reuses its detectors of
Lemma~\ref{lem:detector} to remove irrelevant messages from the
system.

>>>>

\begin{proof}
  %% knowing that b_p delta is sufficient to say that processes will
  %% receive delta or detect blocking
  Assuming a chain of delivery $\pi_{xz} = [x, \ldots, z]$ of
  $\alpha_x$ with $\alpha_x = \last D_z$ and $d_x(\delta_x$), we must
  prove that whatever the possible states of \nodes that belong to
  this chain, it either leads to outcome (\ref{lem:detectorA}) or
  (\ref{lem:detectorB}).
  \begin{asparadesc}
  \item [(I) Same last partition:] $\forall y \in \pi_{xz}: \alpha_x =
    \last D_y$, with forwarding of staleness
    (Definition~\ref{def:fs}), $f_{\pi_{xz}[k]}(\delta_x) \implies
    \eventually d_{\pi_{xz}[k+1]}(\delta_x)$ except if $\delta_x \in
    D_{\pi_{xz}[k+1]}$ which means that $\pi_{xz}[k+1]$ already
    delivered and forwarded $\delta_x$ following another such a chain
    of delivery. Therefore, $z$ eventually receives, hence delivers
    $\delta_x$ (outcome \ref{lem:detectorA}).

  \item[(II) Different last partitions:] $\exists y \in \pi_{xz}
    \setminus \{x, z\}: \alpha_s = \last D_y$, by (I), the staleness
    notification reaches the first of such a \process $y =
    \pi_{xz}[k]$.  Forwarding of staleness (Definition~\ref{def:fs})
    states that the forwarding stops when $y$ already delivered
    $\delta_x$ ($\delta_x \in D_y$ covered by (I)) or delivered a
    better message $\alpha_s = \last D_y$.  With $y'=\pi_{xz}[k+1]$,
    forwarding of best (Definition~\ref{def:fbplus}) implies three
    possible results:
  \begin{asparadesc}
  \item [(i) $y'$ equivalent to $y$:] $\alpha_{s'} = \last D_{y'}$
    with $\alpha_{s'} \neq \alpha_x$ hence $y' \iff y$ which leaves
    two possible results as follows:
  \item [(ii) $y'$ in $P_x$ from another parent:] $r_{y'}(\alpha_s)
    \wedge \alpha_x = \last D_{y'} \wedge \alpha_x < \alpha_s$ which
    means that a shorter path of delivery $\pi_{xz}'$ exists that
    either forwards appropriate staleness notifications (covered by
    (I)) or detects its possible blocking as follows:
  \item [(iii) $y'$ in $P_x$ with $y$ as parent] but does not deliver
    the $y$'s last partition for it already delivered the
    corresponding staleness notification: $r_{y'}(\alpha_s) \wedge
    \delta_s \in D_{y'} \wedge \alpha_x = \last D_{y'}$ which detects
    a possible blocking of $\delta_x$ (outcome
    (\ref{lem:detectorB})). Without global knowledge, $y'$ assumes it
    belongs to the shortest and only path of delivery of $\alpha_x$
    % ($\pi_{xy'} \subseteq \Pi_{xz}$)
    thus it cannot further delegate the detection to another \process.
  \end{asparadesc}
  \end{asparadesc}
  \vspace{-1.5em} %% ugly vspace but necessary for enumerate+qed
  %
  %% (i) If every \process (such as $a$) starting from the source
  %% delivers and forwards a removal notification $\delta_X$ when their
  %% last delivery is $\alpha_X$, then every such \process eventually
  %% delivers the removal notification $\delta_X$ except \processes (such
  %% as $b$) that delivered $\alpha_X$ from a \process that delivered a
  %% message from another partition $\alpha_Y$ since then.
  %
  %% These exceptions eventually receive and deliver $\alpha_Y$ since
  %% $\alpha_Y$ is better than $\alpha_X$ through this path, except if
  %% they already received and delivered the removal notification of
  %% $P_Y$ (such as $c$). These \processes may never receive hence
  %% deliver $\delta_X$, and may never receive hence deliver a better
  %% message than $\alpha_X$. They need an additional mechanism to
  %% eventually purge $\alpha_X$ that cannot rely on the eventual purging
  %% of $P_Y$ at \processes like $b$, to avoid deadlocks.
  %
  %% (ii) Assuming that every \process keeps an history of its past
  %% deliveries, \processes (such as $c$) can detect the inconsistency,
  %% since they receive from their parent an already deleted partition
  %% $P_Y$. This notification means that the parent discards any
  %% $\delta_Z$ with $\alpha_Z^z < \alpha_Y^y$, and most importantly, if
  %% $\delta_X$ exists, it discards it. To ensure the purge of stale
  %% messages, a detecting \process must assume the worst case that such
  %% $\delta_X$ exists, and send another kind of message, noted $\Delta$,
  %% that notify the possible removal of $P_X$.
  %
  %% (iii) A \process (such as $d$) whose last delivery is $\alpha_X$,
  %% but whose parent is neither inconsistent (like $b$) nor receiving
  %% $\delta_X$ (like $a$), eventually receives $\Delta$ from a detecting
  %% \process, either directly or transitively, for such a child \process
  %% ($d$) trusts the possible removal of $P_X$ by delivering and
  %% forwarding such $\Delta$. $\Delta$ suffers from identical blocking
  %% conditions (between $b$ and $c$) than $\delta$, leading to the same
  %% solutions of detection and propagation of $\Delta$. Eventually,
  %% every \process whose last delivery is $\alpha_X$ receives and
  %% delivers either a better $\alpha_Z$, or $\delta_X$, or $\Delta_X$.
\end{proof}

As a second step, detecting \processes must proactively purge the
system from their forwarded notifications. For instance in
Figure~\ref{fig:problemF}, $c$ detects the blocking of the staleness
notification $\delta_a$. It could broadcast $\delta_a$ in order to
acknowledge the staleness of its own $\alpha_a^3$ and inform
neighboring \processes that delivered it as well. Unfortunately, there
exists false positive in the blocking detection.
Lemma~\ref{lem:detector}'s proof shows that the detection at $y'$ does
not depend on $\delta_x$. Therefore $\delta_x$ may not exist but $y'$
still receives $\alpha_s$ after the delivery of $\delta_s$ from the
path it received and delivered
$\alpha_x$. Figure~\ref{fig:undoproblem} highlights this behavior. In
Figure~\ref{fig:undoproblemB}, $a$ does not delete its partition. In
Figure~\ref{fig:undoproblemF} and because of scoped broadcast, $c$
received the same series of messages as in
Figure~\ref{fig:problemF}. Yet, $c$ must assume the existence of
$\delta_a$ and act accordingly by forcing the staleness of its best
delivered message and disseminate this information to its neighbors.
To avoid flooding the system with false positive staleness
notifications, we reduce the scope of staleness notifications by
including only downstream \processes. In
Figure~\ref{fig:undoproblemF}, false positives would generate traffic
at $c$ and its children if it had any, without impacting $a$ and $b$.

@@

Algorithm~\ref{algo:ascast} shows the instructions of our
implementation called \NAME that enables dynamic consistent
partitioning. It implements forwarding of best (see
Line~\ref{line:fb}) and echos (see Line~\ref{line:echo}). To implement
forwarding of staleness downstream as stated in
Definition~\ref{def:fsplus}:
\begin{inparaenum}[(A)]
\item each \process maintains a vector of versions that associates the
  respective known local counter of each known source, or has-been
  source. It constitutes a summary of known progress of other
  \processes; 
\item each notification message $\alpha$ carries the list of
  identifier and counter of each \node that forwarded it. In the worst
  case, both these structures grow linearly with the number of
  \processes in the system $\mathcal{O}(V)$. Nevertheless, following
  the principles of scoped broadcast, we expect that \processes only
  acknowledge a small subset of sources and messages;
\item each staleness notifications $\delta$ only carry the identifier
  and counter of the \process~--~source or detector~--~that generated
  it. Only downstream \processes may deliver such message, since they
  carry the identifier of the generator. 
\end{inparaenum}
To implement the detection as stated in Lemma~\ref{lem:detector}, each
\process only requires to know the direct parent of its best delivered
message which is already included in the piggybacked path of this
message. Receiving a message known to be stale from this parent
triggers the generation of staleness notifications that can only be
delivered by downstream \processes (see Line~\ref{line:detect}).

@@

By reusing \NAME's default behavior of echos and downstream staleness,
Algorithm~\ref{algo:edges} enables dynamic consistent partitioning
even in dynamic networks subject to physical partitioning where
\processes can join, leave, or crash at any time. Adding a \process is
equivalent to add as many edges as necessary. Since adding an edge may
improve shortest paths, the protocol triggers a competition using
echos of Definition~\ref{def:fbplus}.  Removing a \process is
equivalent to removing all its edges. Removing an edge between two
\processes may invalidate the shortest path of one of involved
\processes plus downstream \processes, or impair the propagation of
staleness notifications. Therefore, the protocol reuses its detectors
of Lemma~\ref{lem:detector} to remove irrelevant messages from the
system.
